Analysis paper

### 2021-06-02 analysis/analysis.Rmd
update plots with th new ensemble

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(knitr.kable.NA = '')
library(covid.german.forecasts)
library(here)
library(magrittr)
library(scoringutils)
library(knitr)
library(kableExtra)
library(data.table)
library(ggplot2)
library(ggridges)
library(RColorBrewer)
library(patchwork)
library(stringr)
```

@@

```{r truth-vs-forecasts, fig.height=12, fig.width=15}
# create helper function for the plot ------------------------------------------
truth_and_pred_plot <- function(df) {
  p <- df %>%
  ggplot(aes(y = true_value, x = target_end_date)) + 
  geom_linerange(aes(ymin = `0.25`, ymax = `0.75`, color = model), 
                 size = 1.5,
                 alpha = 1,
                 position = position_dodge(width = 6)) + 
  labs(y = "value", x = "Date") + 
  geom_point(aes(shape = classification)) + 
  geom_line() + 
  theme_light() + 
  facet_wrap(~ location_target, scales = "free", ncol = 2) + 
  theme(legend.position = "bottom") + 
  theme(panel.background = element_rect(fill = "#e1f0fe")) + 
  scale_x_date(limits = c(min(forecast_dates$full_hub_period) + 14, 
                          max(forecast_dates$full_hub_period)))
  return(p)
}

@@

df <- unfiltered_data[quantile %in% c(0.025, 0.25, 0.75, 0.975)] %>%
  dcast(... ~ quantile, value.var = "prediction")
df <- merge(
  df[, c("classification", "speed") := NULL], 
  copy(epitrend)[, forecast_date := NULL], 
  by = c("target_end_date", "location_name", "target_type")
)

@@

for (horizon in 1:4) {
  # plot for cases
  truth_vs_forecast_cases <- df[grepl(horizon, target)] %>%
  dplyr::filter(target_type == "case") %>%
  truth_and_pred_plot() + 
  scale_color_manual(values = brewer.pal(name = "Accent", n = 5)[2:5])

@@

  # plot for deaths
  truth_vs_forecast_deaths <- df[grepl(horizon, target)] %>%
  dplyr::filter(target_type == "death") %>%
  truth_and_pred_plot()  + 
  scale_color_manual(values = brewer.pal(name = "Accent", n = 5)[1:5]) + 
  annotate('rect', xmin = min(forecast_dates$full_hub_period) + 14,
                xmax = max(forecast_dates$death_not_scored) + 14,
                ymin = 0,
                ymax = Inf, fill = "grey10", alpha = 0.1)

@@

```{r scores-over-time, fig.height=10, fig.width=15}
# score forecasts --------------------------------------------------------------
# obtain scores and then do pairwise comparisons between models
summarise_by <- c("model", "horizon", "location_name", "target_type", 
                   "forecast_date", "target_end_date", "location_target", 
                   "classification")
scores_target_forecastdate <- eval_forecasts(
  data = unfiltered_data,
  summarise_by = summarise_by
)

@@

# create helper function to plot forecasts -------------------------------------
scores_over_time_plot <- function(scores, y = "mean_scores_ratio") {
  p <- scores %>%
  ggplot(aes(x = as.Date(target_end_date), y = get(y))) + 
  geom_line(aes(colour = model)) + 
  geom_point(aes(shape = classification, colour = model)) + 
  theme_light() + 
  facet_wrap( ~ location_target, ncol = 2, scales = "free_y") +
  theme(legend.position = "bottom") +
  labs(y = "WIS relative to baseline", x = "Date") + 
  # geom_rect(aes(xmin = min(forecast_dates$christmas) + 14,
  #               xmax = max(forecast_dates$christmas) + 14,
  #               ymin = 0,
  #               ymax = Inf), fill = "lightgrey", alpha = 0.002) +
  theme(panel.background = element_rect(fill = "#e1f0fe")) +
  scale_y_log10() + 
  scale_x_date(limits = c(min(forecast_dates$full_hub_period) + 14, 
                          max(forecast_dates$full_hub_period)))
  return(p)
}

# create plots for all horizons ------------------------------------------------
for (i in 1:4) {
  scores_over_time_case <- pairwise[horizon == i & target_type == "case"] %>%
    scores_over_time_plot() + 
    scale_color_manual(values = brewer.pal(name = "Accent", n = 5)[2:5])

@@

  scores_over_time_deaths <- pairwise[horizon == i & target_type == "death"] %>%
    scores_over_time_plot() + 
    scale_color_brewer(palette = "Accent") + 
    annotate('rect', xmin = min(forecast_dates$full_hub_period) + 14,
             xmax = max(forecast_dates$death_not_scored) + 14,
             ymin = 0,
             ymax = Inf, fill = "grey10", alpha = 0.1)

@@

horizons <- 1:4

@@

  ggsave(here("analysis", "plots", paste0("figure-forecasts-", horizon, ".png")))
}

@@

*Figure 1. A and C: Visualisation of 2 week ahead forecasts against the true observed values. The shape indicates whether there has been a monotonic increase or decrease over the last two weeks, or an unclear trend. Forecasts made during the Christmas period (from 2020-12-19 until 2021-01-02) are grayed out. B and D: Corresponding scaled relative skill scores for the forecasts shown on the left. Scaled relative skill scores can be thought of as 'improvement over the baseline model' (see Methods for details).*

**What we learn from this plot**

-   severe overshooting for Renewal model
-   all models are terrible around Christmas
-   the artifacts around Christmas hit Renewal hardest
-   the Christmas effect is much stronger in Germany than in Poland

@@

```{r }
# create scores ----------------------------------------------------------------
# calculate coverage independently and merge with other scores
coverage <- eval_forecasts(
  filtered_data, 
  summarise_by = c("horizon", "model", "target_type", "range"), 
  compute_relative_skill = TRUE, 
  baseline = "Baseline",
)[range %in% c(50, 90), 
                   .(model, target_type, coverage, range, horizon)] %>%
  dcast(formula = ... ~ range, value.var = "coverage")

df <- eval_forecasts(
  filtered_data, 
  summarise_by = c("model", "horizon", "target_type"), 
  compute_relative_skill = TRUE, 
  baseline = "Baseline"
) %>%
  merge(coverage, all.x = TRUE, by = c("model", "horizon", "target_type"))

@@

*Table 1: Scores for 2 week ahead forecasts (excluding the 21st and 28th of December 2020, cut to three significant digits and rounded). Skill is the scaled relative skill, a measure of relative performance with respect to the baseline model. Sharpness, overprediction and underprediction together some up to the weighted interval score (WIS). Bias (between -1 and 1) represents the general average tendency of a model to over- or underpredict. 50% and 90%-coverage are the percentage of observed values that fell within the 50% and 90% prediction intervals of a model.*

@@

df <- eval_forecasts(
  filtered_data, 
  metric = "interval_score",
  summarise_by = c("model", "horizon", "target_type"), 
  sd = TRUE, 
  quantile = 0.5
)

@@

df <- eval_forecasts(
  filtered_data, 
  summarise_by = c("model", "horizon", "target_type")
)

@@

coverage <- eval_forecasts(
  filtered_data, 
  summarise_by = c("horizon", "model", "target_type", "target_phase", "range"), 
)[range %in% c(50, 90), 
                   .(model, target_type, coverage, range, target_phase, horizon)] %>%
  dcast(formula = ... ~ range, value.var = "coverage")

df <- eval_forecasts(
  filtered_data, 
  summarise_by = c("model", "horizon", "target_type", "target_phase"), 
  metrics = c("interval_score", "coverage_deviation"),
  sd = TRUE, 
  quantiles = 0.5
) %>%
  merge(coverage, all.x = TRUE, by = c("model", "horizon", "target_phase", "target_type"))

@@

```{r}
scores <- eval_forecasts(
  filtered_data[grepl("2", target)], 
  summarise_by = c("model", "target_type", "location_name", "forecast_date", 
                   "target_phase",
                   "location_target", "classification"), 
  compute_relative_skill = TRUE, 
  baseline = "Baseline"
)

@@

# scores[, forecast_date := as.character(forecast_date)]
# scores <-  merge(scores, epitrend, 
#             by = c("forecast_date", "location_name", "target_type"))

```{r}
scores[, target_type := paste0(str_to_title(target_type), "s overall")]

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(knitr.kable.NA = '')
library(covid.german.forecasts)
library(here)
library(magrittr)
library(scoringutils)
library(knitr)
library(kableExtra)
library(data.table)
library(ggplot2)
library(ggridges)
library(RColorBrewer)
library(patchwork)
library(stringr)

@@

```{r truth-vs-forecasts, fig.height=12, fig.width=15}
# create helper function for the plot ------------------------------------------
truth_and_pred_plot <- function(df) {
  p <- df %>%
  ggplot(aes(y = true_value, x = target_end_date)) + 
  geom_linerange(aes(ymin = `0.25`, ymax = `0.75`, color = model), 
                 size = 1.5,
                 alpha = 1,
                 position = position_dodge(width = 6)) + 
  geom_point(aes(shape = classification)) + 
  geom_line() + 
  theme_light() + 
  facet_wrap(~ location_target, scales = "free", ncol = 2) + 
  theme(legend.position = "bottom") + 
  theme(panel.background = element_rect(fill = "#e1f0fe")) + 
  scale_x_date(limits = c(min(forecast_dates$full_hub_period) + 14, 
                          max(forecast_dates$full_hub_period)))
  return(p)
}

@@

df <- unfiltered_data[model %in% regular_models &
                        quantile %in% c(0.025, 0.25, 0.75, 0.975)] %>%
  dcast(... ~ quantile, value.var = "prediction")
df <- merge(
  df[, c("classification", "speed") := NULL], 
  copy(epitrend)[, forecast_date := NULL], 
  by = c("target_end_date", "location_name", "target_type")
)

@@

for (horizon in 1:4) {
  # plot for cases
  truth_vs_forecast_cases <- df[grepl(horizon, target)] %>%
    dplyr::filter(target_type == "case") %>%
    truth_and_pred_plot() + 
    scale_color_manual(values = brewer.pal(name = "Accent", n = 5)[-2]) + 
    labs(y = "Cases", x = "")

@@

  # plot for deaths
  truth_vs_forecast_deaths <- df[grepl(horizon, target)] %>%
  dplyr::filter(target_type == "death") %>%
  truth_and_pred_plot()  + 
  scale_color_manual(values = brewer.pal(name = "Accent", n = 5)[1:5]) + 
  annotate('rect', xmin = min(forecast_dates$full_hub_period) + 14,
                xmax = max(forecast_dates$death_not_scored) + 14,
                ymin = 0,
                ymax = Inf, fill = "grey10", alpha = 0.1) + 
    labs(y = "Deaths", x = "")

@@

```{r scores-over-time, fig.height=10, fig.width=15}
# score forecasts --------------------------------------------------------------
# obtain scores and then do pairwise comparisons between models
summarise_by <- c("model", "horizon", "location_name", "target_type", 
                   "forecast_date", "target_end_date", "location_target", 
                   "classification")
scores_target_forecastdate <- eval_forecasts(
  data = unfiltered_data[model %in% regular_models],
  summarise_by = summarise_by
)

@@

# create helper function to plot forecasts -------------------------------------
scores_over_time_plot <- function(scores, y = "mean_scores_ratio") {
  p <- scores %>%
  ggplot(aes(x = as.Date(target_end_date), y = get(y))) + 
  geom_line(aes(colour = model)) + 
  geom_point(aes(shape = classification, colour = model)) + 
  theme_light() + 
  facet_wrap( ~ location_target, ncol = 2, scales = "free_y") +
  theme(legend.position = "bottom") +
  # geom_rect(aes(xmin = min(forecast_dates$christmas) + 14,
  #               xmax = max(forecast_dates$christmas) + 14,
  #               ymin = 0,
  #               ymax = Inf), fill = "lightgrey", alpha = 0.002) +
  theme(panel.background = element_rect(fill = "#e1f0fe")) +
  scale_y_log10() + 
  scale_x_date(limits = c(min(forecast_dates$full_hub_period) + 14, 
                          max(forecast_dates$full_hub_period)))
  return(p)
}

# create plots for all horizons ------------------------------------------------
for (i in 1:4) {
  scores_over_time_case <- pairwise[horizon == i & target_type == "case"] %>%
    scores_over_time_plot() + 
    scale_color_manual(values = brewer.pal(name = "Accent", n = 5)[-2]) + 
    labs(y = "WIS rel. to baseline", x = "")

@@

  scores_over_time_deaths <- pairwise[horizon == i & target_type == "death"] %>%
    scores_over_time_plot() + 
    scale_color_brewer(palette = "Accent") + 
    annotate('rect', xmin = min(forecast_dates$full_hub_period) + 14,
             xmax = max(forecast_dates$death_not_scored) + 14,
             ymin = 0,
             ymax = Inf, fill = "grey10", alpha = 0.1) + 
    labs(y = "WIS rel. to baseline", x = "Date")

@@

```{r fig.height=15, fig.width=10}
horizons <- 1:4

@@

  ggsave(here("analysis", "plots", paste0("figure-forecasts-", horizon, ".png")), width = 10, height = 12)
}

@@

@@

```{r }
# create scores ----------------------------------------------------------------
# calculate coverage independently and merge with other scores
coverage <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("horizon", "model", "target_type", "range"), 
)[range %in% c(50, 90), 
                   .(model, target_type, coverage, range, horizon)] %>%
  dcast(formula = ... ~ range, value.var = "coverage")

df <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("model", "horizon", "target_type"), 
  compute_relative_skill = TRUE, 
  baseline = "Baseline"
) %>%
  merge(coverage, all.x = TRUE, by = c("model", "horizon", "target_type"))

@@

### Create table for all median ensembles

```{r }
# create scores ----------------------------------------------------------------
# calculate coverage independently and merge with other scores
coverage <- eval_forecasts(
  filtered_data[model %in% ensemble_models[!grepl("mean", ensemble_models)]], 
  summarise_by = c("horizon", "model", "target_type", "range"), 
)[range %in% c(50, 90), 
                   .(model, target_type, coverage, range, horizon)] %>%
  dcast(formula = ... ~ range, value.var = "coverage")

df <- eval_forecasts(
  filtered_data[model %in% ensemble_models[!grepl("mean", ensemble_models)]], 
  summarise_by = c("model", "horizon", "target_type"), 
  compute_relative_skill = TRUE, 
  baseline = "Baseline"
) %>%
  merge(coverage, all.x = TRUE, by = c("model", "horizon", "target_type"))

# tables for 1-2 weeks ahead and 3-4 weeks ahdead
table <- make_score_table(df[horizon <= 2]) 
save_kable(table, file = here("analysis", "plots", "table_median-ensemble_scores_2_ahead.png"))

table <- make_score_table(df[horizon > 2]) 
save_kable(table, file = here("analysis", "plots", "table_median-ensemble_scores_4_ahead.png"))

### Create table for all mean ensembles

```{r }
# create scores ----------------------------------------------------------------
# calculate coverage independently and merge with other scores
coverage <- eval_forecasts(
  filtered_data[model %in% ensemble_models[grepl("mean", ensemble_models)]], 
  summarise_by = c("horizon", "model", "target_type", "range"), 
)[range %in% c(50, 90), 
                   .(model, target_type, coverage, range, horizon)] %>%
  dcast(formula = ... ~ range, value.var = "coverage")

df <- eval_forecasts(
  filtered_data[model %in% ensemble_models[grepl("mean", ensemble_models)]], 
  summarise_by = c("model", "horizon", "target_type"), 
  compute_relative_skill = TRUE, 
  baseline = "Baseline"
) %>%
  merge(coverage, all.x = TRUE, by = c("model", "horizon", "target_type"))

# tables for 1-2 weeks ahead and 3-4 weeks ahdead
table <- make_score_table(df[horizon <= 2]) 
save_kable(table, file = here("analysis", "plots", "table_mean-ensemble_scores_2_ahead.png"))

table <- make_score_table(df[horizon > 2]) 
save_kable(table, file = here("analysis", "plots", "table_mean-ensemble_scores_4_ahead.png"))

@@

df <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  metric = "interval_score",
  summarise_by = c("model", "horizon", "target_type"), 
  sd = TRUE, 
  quantile = 0.5
)

@@

df <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("model", "horizon", "target_type")
)

@@

coverage <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("horizon", "model", "target_type", "target_phase", "range"), 
)[range %in% c(50, 90), 
                   .(model, target_type, coverage, range, target_phase, horizon)] %>%
  dcast(formula = ... ~ range, value.var = "coverage")

df <- eval_forecasts(
  filtered_data[model %in% regular_models], 
  summarise_by = c("model", "horizon", "target_type", "target_phase"), 
  metrics = c("interval_score", "coverage_deviation"),
  sd = TRUE, 
  quantiles = 0.5
) %>%
  merge(coverage, all.x = TRUE, by = c("model", "horizon", "target_phase", "target_type"))

@@

```{r}
scores <- eval_forecasts(
  filtered_data[(model %in% regular_models) & grepl("2", target)], 
  summarise_by = c("model", "target_type", "location_name", "forecast_date", 
                   "target_phase",
                   "location_target", "classification"),
  compute_relative_skill = TRUE, 
  baseline = "Baseline"
)

@@

scores[, target_type := paste0(str_to_title(target_type), "s overall")]
